Trans-Dimensional Inference for Gravitational Lens Substructures

### 2015-06-24 Papers/ms.tex
Monte carlo, fixed mistake
More empirical measure talk

The {\it empirical measure} of the substructure positions is a function that
takes the substructure positions $(x_i^{\rm lens}, y_i^{\rm lens})$ and
produces a ``density function'' over two dimensions, composed of delta
functions at the positions themselves:

\begin{eqnarray}
\left\{
(x_i^{\rm lens}, y_i^{\rm lens})
\right\}_{i=1}^{N_{\rm lens}}
\Longrightarrow
\sum_{i=1}^{N_{\rm lens}}
\delta\left(x - x_i^{\rm lens}, y - y_i^{\rm lens}\right).
\end{eqnarray}

The empirical measure, being a function of the actual substructure positions,
is not available to us since we do not know those positions with certainty.
However, we have samples from the posterior distribution for those positions,
and can use these (trivially) to create samples from the posterior distribution
for the empirical measure. We can also summarise this posterior, for example,
by taking its expected value.

The posterior expected value of the empirical measure is:
\begin{eqnarray}
\int p(\boldsymbol{\theta} | \boldsymbol{D})
\sum_{i=1}^{N_{\rm lens}}
\delta\left(x - x_i^{\rm lens}, y_i^{\rm lens} - y_i^{\rm lens}\right)
\, d\boldsymbol{\theta}
\end{eqnarray}
where $\boldsymbol{\theta}$ denotes all parameters and hyperparameters
(including the $N$s) and ``$\int \,d\boldsymbol{\theta}$'' is an integral and summation over the entire parameter space.

>>>>

The {\it empirical measure} of the substructure positions is a function that
takes the substructure positions $(x_i^{\rm lens}, y_i^{\rm lens})$ and
produces a ``density function'' over two dimensions, composed of delta
functions at the positions themselves:
\begin{eqnarray}
\left\{
(x_i^{\rm lens}, y_i^{\rm lens})
\right\}_{i=1}^{N_{\rm lens}}
\Longrightarrow
\sum_{i=1}^{N_{\rm lens}}
\delta\left(x - x_i^{\rm lens}, y - y_i^{\rm lens}\right).
\end{eqnarray}

Intuitively, the empirical measure is a mathematical object that is like
an ``infinite resolution histogram'', in this case a two dimensional histogram,
of the substructure positions.
Being a function of the actual substructure positions, the empirical measure
is not available to us since we do not know those positions with certainty.
However, we have samples from the posterior distribution for those positions,
and can use these (trivially) to create samples from the posterior distribution
for the empirical measure. We can also summarise this posterior, for example,
by taking its expected value.

The posterior expected value of the empirical measure is:
\begin{eqnarray}
\int p(\boldsymbol{\theta} | \boldsymbol{D})
\sum_{i=1}^{N_{\rm lens}}
\delta^2\left(x - x_i^{\rm lens}, y_i^{\rm lens} - y_i^{\rm lens}\right)
\, d\boldsymbol{\theta}
\end{eqnarray}
where $\boldsymbol{\theta}$ denotes all parameters and hyperparameters
(including the $N$s) and ``$\int \,d\boldsymbol{\theta}$'' is an integral and summation over the entire parameter space.

Since we can approximate posterior expectations using Monte Carlo, we can
obtain the expected value of the empirical measure using:
\begin{eqnarray}
\sum_{k=1}^n
\sum_{i=1}^{N_{\rm lens}}
\delta^2\left(x - x_i^{\rm lens}, y_i^{\rm lens} - y_i^{\rm lens}\right)
\, d\boldsymbol{\theta}
\end{eqnarray}
where $n$ is the number of posterior samples. The resulting function is
an ``image'' with a point mass wherever a substructure occurred. For
visualisation purposes the image can be blurred, or calculated at a lower
resolution by replacing the Dirac-delta function with a discrete version
which returns a nonzero constant
if a substructure appears in a pixel or zero otherwise.
